Hourly Assimilation of Different Sources of Observations Including Satellite Radiances in a  Mesoscale Convective System Case During RELAMPAGO campaign

### 2022-07-26 analysis/paper/paper_ana_20181122.Rmd
Posible weather field y otros

For DA methods to be successful, observing networks with sufficient temporal and spatial resolution capable of capturing mesoscale variability should be used.

In that regard, several authors investigated the impact of assimilating surface weather data (e.g. @wheatley2010, @ha2014, @chang2017, @bae2022, @banos2021, @maejima2019, and @chen2016) using different assimilation methodologies. Most of these studies reported the beneficial impacts of assimilating temperature and dew point observations upon the planetary boundary layer structure and the location and timing of precipitating systems.  @sobash2015 showed using a mesoscale DA system, that the positive impact upon convection initiation and the short range precipitation forecast is achieved if data is assimilated frequently (in the order of minutes, rather than in the order of hours).  More recently, @gasperoni2018 evaluated the impact of assimilating observations produced by private weather station networks which are not incorporated in the operational analysis. They found a positive effect of these observations upon the initiation of deep moist convection. This result is particularly important for data sparse regions such as Southern South America, where operational networks are not dense enough to capture mesoscale details.

@@

While the assimilation of radiance observations into global models is well established [@eyre2020], the direct assimilation of radiance data into regional models, however, still remains a challenge due to the sparse data coverage, bias correction, and the relatively low model tops used for this application. @bao2015 studied the impact of assimilating microwave and infrared radiance data on temperature and humidity forecasts over the western USA and found a reduction in the temperature bias at low and mid-levels as a result of the microwave observations but an opposite effect for infrared data. More recently, @zhu2019 studied the impact of assimilating satellite radiance data within a frequently updated regional system and showed an improvement for all variables, in particular for relative humidity at upper levels. @wang2021 studied the impact of assimilating radiances in the high-resolution Copernicus European Regional Reanalysis. They reported that satellite radiance observations had a neutral impact on the analyses of geopotential height in the lower troposphere, while a slightly negative impact on the upper troposphere and the stratosphere. They also observed similar results for 3-h forecasts initialized from the analysis but a positive impact on 12 and 24 -h forecasts. Given these mixed results, there is still room to analyze the utility of assimilating radiance observations in a limited-area DA system over land. Moreover, to the best of our knowledge, there are no studies related to the direct assimilation of radiance observations over South America.

@@

## Case Study

On Nov 22, 2018 a cold front crossed the center of Argentina triggering isolated convective cells that rapidly grew upscale into an exceptionally large MCS. To the north of the region, a warm front contributed to the development of isolated multicells that ultimately grew and merged with the MCS. 
The MCS traveled approximately 2500 km from south to north, dissipating over Paraguay and Southern Brazil after 42 hours.

@@

To reduce the effect of spurious correlations in the estimation of error covariances, we use a horizontal localization radius of 180 km and a vertical localization radius of 0.4 (in log pressure coordinates) as in @dillon2021 for all types of observations. 
A relaxation-to-prior inflation [@whitaker2012] is applied with an inflation parameter $\alpha=0.9$ to mitigate the impact of sampling errors and to consider model errors not accounted for by the multi-model ensemble approach.

@@

The impacts on the wind components are shown in Figure \@ref(fig:UV-diff), along with the corresponding averaged wind component in the experiment with the largest number of assimilated observations (for example, Figure \@ref(fig:UV-diff)a shows the zonal wind difference between AWS and CONV and the zonal wind for AWS). The assimilation of AWS produces a more easterly wind and a less northerly wind at low levels during the first two days of analysis (Figures \@ref(fig:UV-diff)a,b). There is a diurnal cycle in the impact of surface weather stations on the meridional velocity  (Figure  \@ref(fig:UV-diff)d) with a stronger reduction of the northerly wind during night hours. This indicates that surface observations are reducing the intensity of the low level jet present in the pre-convective environment. After 18 UTC Nov 22, the opposite effect is observed when the MCS is moving through the domain. After the initiation of the convective cells, the systematic impact on the wind field is larger at mid and upper levels (Figures  \@ref(fig:UV-diff)d, f). During Nov 22 and 23 the impact of assimilating AWS observations produces an increase of northerly wind in upper levels. This could be a consequence of a stronger MCS with an increased polar side upper level outflow. Although satellite-derived wind observations produce the largest impact in mid-to-upper levels where the number of observations is largest; the systematic impact is overall smaller than the one produced by assimilating data from AWS  (Figures  \@ref(fig:UV-diff)b, e).

@@

(ref:UV-diff) Difference between analysis ensemble mean experiments a) and d) AWS-CONV, b) and e) SATWND-AWS, and c) and f) RAD-SATWND for the spatially averaged vertical profiles of u wind (a, b, and c, in $m\ s^{-1}$) and v wind (d, e, and f in $m\ s^{-1}$) calculated over the inner domain (red box in Figure \@ref(fig:dominio)a) for each analysis cycle. Black contours correspond to u wind for (a) AWS, (b) SATWND, and (c) RAD and v wind for (d) AWS, (e) SATWND, and (f) RAD. Dashed contours correspond to negative v wind.

>>>>

For DA methods to be successful, observing networks with sufficient temporal and spatial resolution capable of capturing mesoscale variability should be used. In that regard, several authors investigated the impact of assimilating surface weather data (e.g. @wheatley2010, @ha2014, @chang2017, @bae2022, @banos2021, @maejima2019, and @chen2016) using different assimilation methodologies. Most of these studies reported the beneficial impacts of assimilating temperature and dew point observations upon the planetary boundary layer structure and the location and timing of precipitating systems.  @sobash2015 showed using a mesoscale DA system, that the positive impact upon convection initiation and the short range precipitation forecast is achieved if data is assimilated frequently (in the order of minutes, rather than in the order of hours).  More recently, @gasperoni2018 evaluated the impact of assimilating observations produced by private weather station networks which are not incorporated in the operational analysis. They found a positive effect of these observations upon the initiation of deep moist convection. This result is particularly important for data sparse regions such as Southern South America, where operational networks are not dense enough to capture mesoscale details.

@@

While the assimilation of radiance observations into global models is well established [@eyre2020], the direct assimilation of radiance data into regional models, however, still remains a challenge due to the sparse data coverage (in the case of polar-orbiting sateññite observations), bias correction, and the relatively low model tops used for this application. @bao2015 studied the impact of assimilating microwave and infrared radiance data on temperature and humidity forecasts over the western USA and found a reduction in the temperature bias at low and mid-levels as a result of the microwave observations but an opposite effect for infrared data. More recently, @zhu2019 studied the impact of assimilating satellite radiance data within a frequently updated regional system and showed an improvement for all variables, in particular for relative humidity at upper levels. @wang2021 studied the impact of assimilating radiances in the high-resolution Copernicus European Regional Reanalysis. They reported that satellite radiance observations had a neutral impact on the analyses of geopotential height in the lower troposphere, while a slightly negative impact on the upper troposphere and the stratosphere. They also observed similar results for 3-h forecasts initialized from the analysis but a positive impact on 12 and 24 -h forecasts. Given these mixed results, there is still room to analyze the utility of assimilating radiance observations in a limited-area DA system over land. Moreover, to the best of our knowledge, there are no studies related to the direct assimilation of radiance observations over South America.

@@

## Case overview

On Nov 22, 2018 a cold front crossed the center of Argentina triggering isolated convective cells (Figure \@ref(fig:case)a) that rapidly grew upscale into an exceptionally large MCS. During that day several surface stations reported lightning, gustfronts and heavy rain. To the north of the region, a warm front contributed to the development of isolated multicells that ultimately grew and merged with the MCS (Figure \@ref(fig:case)c). 
The MCS traveled approximately 2500 km from south to north, dissipating over Paraguay and Southern Brazil after 42 hours.

(ref:case) GOES-16 infrared imagery corresponding to channel 13 (13 mu) for a) 00 UTC, b) 12 UTC Nov 22, and c) 00 UTC Nov 23.

```{r case, fig.cap="(ref:case)"}

campos <- ReadNetCDF(here("analysis/data/derived_data/reanalysis/geopotential.nc")) %>% 
  dcast(time + longitude + latitude ~ level) %>% 
  .[order(time, -latitude, longitude)] %>% 
  .[, ":="(msl = ReadNetCDF(here("analysis/data/derived_data/reanalysis/pressure-pw.nc"), vars = "msl", out = "vector")[[1]],
           cape = ReadNetCDF(here("analysis/data/derived_data/reanalysis/cape-pw.nc"), vars = "cape", out = "vector")[[1]],
           tcw = ReadNetCDF(here("analysis/data/derived_data/reanalysis/cape-pw.nc"), vars = "tcw", out = "vector")[[1]],
           tcwv = ReadNetCDF(here("analysis/data/derived_data/reanalysis/cape-pw.nc"), vars = "tcwv", out = "vector")[[1]],
           td = ReadNetCDF(here("analysis/data/derived_data/reanalysis/td2m.nc"), vars = "d2m", out = "vector")[[1]])]

campos <-  ReadNetCDF(here("analysis/data/derived_data/reanalysis/cape-pw.nc"), vars = "tcw")

campos %>%

  .[time != ymd_h(2018112312)] %>% 
  .[, ":="(espesor = (`500`-`1000`)/100, time = factor(time))] %>%
  ggplot(aes(longitude, latitude)) +
  geom_contour_fill(aes(z = cape, fill = stat(level_d)), breaks = seq(500, 4000, 500)) +
  scale_fill_distiller(super = ScaleDiscretised,
                       name = "CAPE",
                       # palette = "PRGn", direction = 1,
                       palette = "YlOrRd", direction = 1,
                       guide = guide_colorsteps(barwidth = 25, mid = 25,
                                                barheight = 0.5, 
                                                title.position = "left", 
                                                title.vjust = 1,
                                                frame.colour = "black")) +
  geom_mapa() +
  geom_contour2(aes(z = espesor, label = ..level..),
                size = 0.2, color = "darkred", linetype = 2,
                label_size = 2, label.placer = label_placer_n(n = 3)) +
  # label_size = 2, label.placer = label_placer_fraction(frac = 0.75)) +
  geom_contour2(aes(z = msl/100, label = ..level..),
                size = 0.4, color = "black", linetype = 1,
                label_size = 2) +
  cordillera +
  coord_sf(ylim = c(-45, -22), xlim = c(-75, -52)) +
  facet_wrap(vars(time), labeller = labeller(time = c("2018-11-22 00:00:00" = "00 UTC Nov 22",
                                                      "2018-11-22 12:00:00" = "12 UTC Nov 22",
                                                      "2018-11-23 00:00:00" = "00 UTC Nov 23"))) +
  labs(x = NULL, y = NULL, fill = "CAPE" ) +
  theme_minimal() +
  theme(legend.position = "bottom")

@@

To reduce the effect of spurious correlations in the estimation of error covariances, we use a horizontal localization radius of 180 km and a vertical localization radius of 0.4 (in log pressure coordinates) as in @dillon2021 for all types of observations. 
A relaxation-to-prior spread inflation [@whitaker2012] is applied with an inflation parameter $\alpha=0.9$ to mitigate the impact of sampling errors and to consider model errors not accounted for by the multi-model ensemble approach.

@@

The impacts on the wind components are shown in Figure \@ref(fig:UV-diff), along with the corresponding averaged wind component in the experiment with the largest number of assimilated observations (for example, Figure \@ref(fig:UV-diff)a shows the zonal wind difference between AWS and CONV and the zonal wind for AWS). The assimilation of AWS produces a more easterly wind and a less northerly wind at low levels during the first two days of analysis (Figures \@ref(fig:UV-diff)a,b). There is a diurnal cycle in the impact of surface weather stations on the meridional velocity  (Figure  \@ref(fig:UV-diff)d) with a stronger reduction of the northerly wind during night hours. This indicates that surface observations are reducing the intensity of the low level jet present in the pre-convective environment. After 18 UTC Nov 22, the opposite effect is observed when the MCS is moving through the domain. After the initiation of the convective cells, the systematic impact on the wind field is larger at mid and upper levels (Figures  \@ref(fig:UV-diff)d, f). During Nov 22 and 23 the impact of assimilating AWS observations produces an increase of northerly wind in upper levels. This could be a consequence of a stronger MCS with an increased polar side upper level outflow. Although satellite-derived wind observations produce the largest impact in mid-to-upper levels where the number of observations is largest; the systematic impact is overall smaller than the one produced by assimilating data from AWS  (Figures  \@ref(fig:UV-diff)b, e). The reason of the small impact observed in SATWND could be asociated to the large observation error used for satellite-derived wind observatios.

@@

(ref:UV-diff) Difference between analysis ensemble mean experiments a) and d) AWS-CONV, b) and e) SATWND-AWS, and c) and f) RAD-SATWND for the spatially averaged vertical profiles of u wind (a, b, and c, in $m\ s^{-1}$) and v wind (d, e, and f in $m\ s^{-1}$) calculated over the inner domain (red box in Figure \@ref(fig:dominio)a) for each analysis cycle. Black contours correspond to u wind and aashed contours to negative v wind for (a) AWS, (b) SATWND, and (c) RAD and v wind for (d) AWS, (e) SATWND, and (f) RAD since those experiments are the ones with more assimilated observations in each panel.
